---
title: "Homework 3"
author: Jinghan Zhang
date: October 6, 2020
output: github_document
---

## Problem 1

```{r setup}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Observations are at the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which aisles are the most items from?
```{r aisles and items}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```
 
Let's make a plot

```{r plot aisles and items}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>%
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
Make a table showing the three most popular items in each of the aisles "baking ingredients", "dog food care", and "packaged vegetables fruits". Include the number of times each item is ordered in your table.

```{r}
instacart %>%
  filter(aisle %in% c("baking ingredience", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(rank)
  #knitr::kable()
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
  
```

## Problem 2

```{r setup for Accel data}
library(ggridges)

accel_table = read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() 

accel_table$day = 
  factor(accel_table$day, 
         levels 
            = c("Monday",
                "Tuesday",
                "Wednesday",
                "Thursday",
                "Friday",
                "Saturday",
                "Sunday")
         )

cleaned_accel_table <- accel_table[order(accel_table$week, accel_table$day),] %>%
  mutate(day_type = recode(day,
    "Saturday" = "Weekend",
    "Sunday" = "Weekend",
    .default = "Weekday"
    )) %>%
  relocate(day_type, .after = day)


aggregated <- cleaned_accel_table %>%
  mutate(total_activity = select(cleaned_accel_table, starts_with("activity_")) %>% rowSums()) %>%
  select(week, day_id, day, day_type, total_activity)
  
plot_table <- cleaned_accel_table %>%
  select(day, day_id, starts_with("activity_")) %>%
  gather(minute, activity, activity_1:activity_1440, factor_key = FALSE) %>%
  mutate(minute = as.numeric((str_replace(minute, "activity_", ""))))
  

ggplot(plot_table, aes(x = minute, y = activity))  + 
  geom_point(aes(color = day))
```


```{r setup for noaa data}
library(p8105.datasets)
data("ny_noaa") 

noaa <- ny_noaa %>%
  janitor::clean_names() %>%
  separate(date, c("year", "month", "day"), sep = "-") %>%
  mutate(prcp = as.numeric(prcp)) %>%
  mutate(snow = as.numeric(snow)) %>%
  mutate(snwd = as.numeric(snwd)) %>%
  mutate(tmin = as.numeric(tmin)) %>%
  mutate(tmax = as.numeric(tmax)) %>%
  arrange(year, month, day)

tmax_month <- noaa %>%
  select(-prcp, -snow, -snwd, -tmin) %>%
  spread(day, tmax) %>%
  mutate(avg_tmax = rowMeans(select(., "01":"31"), na.rm = TRUE))

tmax_avg <- tmax_month %>%
  select(id, year, month, avg_tmax)

jan_tmax <- tmax_avg %>%
  filter(month == "01") %>%
  select(id, year, month, avg_tmax)

jul_tmax <- tmax_avg %>%
  filter(month == "07") %>%
  select(id, year, month, avg_tmax)
  
  

```